Explicit Operations on Parametrised Polytopes

### 2018-01-06 dcg-draft.tex
corrected more typos in Theorem 1

\subsection{Properties of Parametrically Convex Maps}\label{ssec:properties:of:p:convex:maps}
%
We begin by introducing an equivalent characterisation of parametric convexity that provides an insight into the geometrical properties of 
set-valued maps satisfying~\eqref{eq:def:parametrically:convex}.
%
This is based on a description of parametric convexity in terms of conditions on the graph~$\mathscr G(\mathcal W)$ of a set-valued map.
%
\begin{definition}\label{def:graph:of:map}
Let $\mathcal W:Y\rightarrow \mathscr P(Z)$ be a continuous set-valued map
such that $\mathcal W(p)$ is convex for all $p\in Y$, then 
%
\begin{equation*}\begin{aligned}
\text{int}(\mathscr G(\mathcal W)) = \{(p,z) \in Y\times Z : \forall \zeta\in\mathbb R^n\;\exists \epsilon>0, \, z+\epsilon \zeta\in \mathcal W(p) \}
\end{aligned}\end{equation*}
%
denotes the \emph{interior} of its graph and
%
\[
  \partial \mathscr G(\mathcal W) = \mathscr G(\mathcal W)\setminus \textup{int}(\mathscr G(\mathcal W))
\]
%
its \emph{boundary};
%
furthermore for any $(p,z)\in\partial\mathscr G(\mathcal W)$ the \emph{normal cone} is defined as 
%
\[
  \mathcal N_{\mathcal W} (p,z) = \{\zeta \in\mathbb R^n: \zeta^T (y - z) \leq 0, \;  \forall y\in \mathcal W(p) \} .
\]
%
\end{definition}
%
\begin{remark}
%
Note that the normal cone in Definition~\ref{def:graph:of:map} is defined in the space of the \emph{set variable} $z\in Y$ rather than \emph{graph variable} $(p,z)\in Y\times Z$.
%
Furthermore, the normal cone $\mathcal N_{\mathcal W} (p,z)$ is the set of normals to all hyperplanes that contain $z$ but no other points in $\mathcal W(p)$.
%
\end{remark}
%
The central idea connecting parametric convexity of a set valued map $\mathcal W$ with properties of its graph $\mathscr G(\mathcal W)$ is stated next.
%
\begin{theorem}\label{thm:p:convexity:graph}
  A map $\mathcal W$ such that $\mathcal W(p)\ni 0$ for all $p\in Y$
  is parametrically convex iff for all $(p_1,z_1), (p_2,z_2)\in\partial\mathscr G(\mathcal W)$
  with
  %$p_1\neq p_2$ and
  $\mathcal N_{\mathcal W}(p_1,z_1)\cap\mathcal N_{\mathcal W}(p_2,z_2)\neq\emptyset$,
%
\begin{equation}\label{eq:graph:def:p:convexity}
\lambda (p_1,z_1) + (1-\lambda) (p_2,z_2) \not\in\textup{int} (\mathscr G(\mathcal W))
\end{equation}
%
holds for all $\lambda\in(0,1)$.
%
\end{theorem}
%
\begin{proof}
  %
  Consider the set $S$, defined for given $\lambda\in (0,1)$ and $p_1,p_2\in Y$ by
  \[
    S = \lambda \mathcal{W}(p_1) \oplus (1-\lambda) \mathcal{W}(p_2) 
      = \bigl\{ z : \lambda z_1 + (1 - \lambda) z_2, \ z_1 \in\mathcal W(p_1), \ z_2 \in\mathcal W(p_2) \bigr\} .
  \]
  % \begin{align*}
  %   S &= \lambda \mathcal{W}(p_1) \oplus (1-\lambda) \mathcal{W}(p_2) \\
  %     &= \bigl\{ z : \lambda z_1 + (1 - \lambda) z_2, \ z_1 \in\mathcal W(p_1), \ z_2 \in\mathcal W(p_2) \bigr\} .
  % \end{align*}
  We first show that $z\in \partial S = S \setminus \textup{int}(S)$ if and only if
  % \begin{equation}\label{eq:Scond}
  %   z = \lambda z_1 + (1 - \lambda) z_2
  %   \quad \text{and} \quad
  %   \mathcal N_{\mathcal W}(p_1,z_1) \cap \mathcal N_{\mathcal W}(p_2,z_2) \neq \emptyset 
  % \end{equation}
  $z = \lambda z_1 + (1 - \lambda) z_2$ for some $z_1\in\partial \mathcal W(p_1)$ and $z_2\in\partial \mathcal W(p_2)$ such that $\mathcal N_{\mathcal W}(p_1,z_1) \cap \mathcal N_{\mathcal W}(p_2,z_2) \neq \emptyset$.
  %

  Specifically, if $z\in \partial S$ then $z\in S$ and there exists $\zeta\in\mathbb R^n$ such that $z + \epsilon \zeta \notin S$ for all $\epsilon > 0$.
  %
  From this it follows that $z = \lambda z_1 + (1-\lambda ) z_2$ for $z_1\in\partial \mathcal W(p_1)$ and $z_2\in\partial \mathcal W(p_2)$ and there exists $n\in \mathbb R^n$ such that for all $\mu \in [0,1]$ we have
  \begin{equation}\label{eq:normal_cone_cond}
    n^T \bigl( \mu \zeta_1 + (1-\mu) \zeta_2 \bigr) \leq 0 
  \end{equation}
  for all $\zeta_1$and $\zeta_2$ such that $n_1^T\zeta_1 \leq 0$ and $n_2^T\zeta_2 \leq 0$ for all $n_1\in\mathcal N_{\mathcal W}(p_1)$ and $n_2\in\mathcal N_{\mathcal W}(p_2)$. Note that these conditions are equivalent to the requirement that the normal cone to $S$ at $z$ is non-empty.
  %
  Clearly $n\in \mathcal N_{\mathcal W}(p_1)$ (since $\mu = 1$ in (\ref{eq:normal_cone_cond}) gives $n^T\zeta_1 \leq 0$) and similarly $n\in \mathcal N_{\mathcal W}(p_2)$ (since $\mu = 0$ in (\ref{eq:normal_cone_cond}) gives $n^T\zeta_2 \leq 0$), and hence $\mathcal N_{\mathcal W}(p_1,z_1) \cap \mathcal N_{\mathcal W}(p_2,z_2) \neq \emptyset$ is necessary for $z\in\partial S$.
  %
  Furthermore $n\in \mathcal N_{\mathcal W}(p_1)\cap \mathcal N_{\mathcal W}(p_2)$ is also sufficient for (\ref{eq:normal_cone_cond}) and hence also $z\in\partial S$.
  %

  The proof is completed by noting that, if $\lambda z_1 + (1-\lambda ) z_2 \in\partial S$, then  (\ref{eq:graph:def:p:convexity}) is equivalent to $\mathcal W\bigl(\lambda p_1 + (1-\lambda) p_2 \bigr) \subseteq S$, and by applying the preceding argument for all $\lambda\in (0,1)$ and all $p_1,p_2 \in Y$.
  \qed
% %
% Assume~\eqref{eq:graph:def:p:convexity} holds for $(p_1,z_1),(p_2,z_2)\in\partial\mathscr G(\mathcal W)$.
% %
% Then the extension of the definition of the Minkowski functional (see e.g.~\cite{Rudin:91}) to the graph $\mathscr G(\mathcal W)$,
% \[
% \mu_{\mathscr G(\mathcal W)} \bigl(
% \mathcal W(p), z \bigr)
% := \min_\mu \{\mu \geq 0 : z \in \mu \mathcal W(p)\},
% \]
% yields $\mu_{\mathscr G(\mathcal W)}\left(\mathcal W(\lambda p_1 + (1-\lambda)p_2),\lambda z_1+(1-\lambda)z_2\right)\geq1$ for all $\lambda\in(0,1)$. Therefore $\lambda z_1 + (1-\lambda) z_2$
% lies either outside the set $\mathcal W(\lambda p_1+(1-\lambda)p_2)$ or on its boundary for all $\lambda\in(0,1)$, and hence the set of all possible interpolation points 
% %
% \[
% \begin{split}
%   \lambda \mathcal W(p_1)\oplus (1-\lambda)\mathcal W(p_2) = \{z : z=\lambda z_1 + (1-\lambda) z_2, z_1\in\mathcal  W(p_1),\, z_2\in\mathcal W(p_2)\}
% \end{split}
% \]
% %
% contains the set $\mathcal W(\lambda p_1 + (1-\lambda)p_2)$ for all $\lambda\in(0,1)$.
% %
% Furthermore, since $\lambda(p_1,z_1)+(1-\lambda)(p_2,z_2)\not\in \text{int}(\mathscr G(\mathcal W))$ and both $(p_1,z_1)$ and $(p_2,z_2)$ lie on the boundary, either $z_1-z_2$ or $z_2-z_1$ is outward facing for both $\mathcal W(p_1)$ and $\mathcal W(p_2)$; therefore either $z_1-z_2$ or $z_2-z_1$ lies in both $\mathcal N_{\mathcal W}(p_1,z_1)$ and $\mathcal N_{\mathcal W}(p_2,z_2)$ and hence also in their intersection.

>>>>

\subsection{Properties of Parametrically Convex Maps}\label{ssec:properties:of:p:convex:maps}
%
We begin by introducing an equivalent characterisation of parametric convexity that provides an insight into the geometrical properties of 
set-valued maps satisfying~\eqref{eq:def:parametrically:convex}.
%
This is based on a description of parametric convexity in terms of conditions on the graph~$\mathscr G(\mathcal W)$ of a set-valued map.
%
\begin{definition}\label{def:graph:of:map}
Let $\mathcal W:Y\rightarrow \mathscr P(Z)$ be a continuous set-valued map
such that $\mathcal W(p)$ is convex for all $p\in Y$, then 
%
\begin{equation*}\begin{aligned}
\text{int}(\mathscr G(\mathcal W)) = \{(p,z) \in Y\times Z : \forall \zeta\in\mathbb R^n\;\exists \epsilon>0, \, z+\epsilon \zeta\in \mathcal W(p) \}
\end{aligned}\end{equation*}
%
denotes the \emph{interior} of its graph and
%
\[
  \partial \mathscr G(\mathcal W) = \mathscr G(\mathcal W)\setminus \textup{int}(\mathscr G(\mathcal W))
\]
%
its \emph{boundary};
%
furthermore for any $(p,z)\in\partial\mathscr G(\mathcal W)$ the \emph{normal cone} is defined as 
%
\[
  \mathcal N_{\mathcal W} (p,z) = \{\zeta \in\mathbb R^n: \zeta^T (y - z) \leq 0, \;  \forall y\in \mathcal W(p) \} .
\]
%
\end{definition}
%
\begin{remark}
%
Note that the normal cone in Definition~\ref{def:graph:of:map} is defined in the space of the \emph{set variable} $z\in Y$ rather than \emph{graph variable} $(p,z)\in Y\times Z$.
%
Furthermore, the normal cone $\mathcal N_{\mathcal W} (p,z)$ is the set of normals to all hyperplanes that contain $z$ but no other points in $\mathcal W(p)$.
%
\end{remark}
%
The central idea connecting parametric convexity of a set valued map $\mathcal W$ with properties of its graph $\mathscr G(\mathcal W)$ is stated next.
%
\begin{theorem}\label{thm:p:convexity:graph}
  A set-valued map $\mathcal W$ satisfying $\mathcal W(p)\ni 0$ for all $p\in Y$
  is parametrically convex iff
%
\begin{equation}\label{eq:graph:def:p:convexity}
\lambda (p_1,z_1) + (1-\lambda) (p_2,z_2) \not\in\textup{int} (\mathscr G(\mathcal W))
\end{equation}
%
holds for all $\lambda\in(0,1)$ and all $(p_1,z_1), (p_2,z_2)\in\partial\mathscr G(\mathcal W)$ such that  $\mathcal N_{\mathcal W}(p_1,z_1)\cap\mathcal N_{\mathcal W}(p_2,z_2)\neq\emptyset$.
%
\end{theorem}
%
\begin{proof}
  %
  Consider the set $S$, defined for given $\lambda\in (0,1)$ and $p_1,p_2\in Y$ by
  \[
    S = \lambda \mathcal{W}(p_1) \oplus (1-\lambda) \mathcal{W}(p_2) 
      = \bigl\{ z : \lambda z_1 + (1 - \lambda) z_2, \ z_1 \in\mathcal W(p_1), \ z_2 \in\mathcal W(p_2) \bigr\} .
  \]
  % \begin{align*}
  %   S &= \lambda \mathcal{W}(p_1) \oplus (1-\lambda) \mathcal{W}(p_2) \\
  %     &= \bigl\{ z : \lambda z_1 + (1 - \lambda) z_2, \ z_1 \in\mathcal W(p_1), \ z_2 \in\mathcal W(p_2) \bigr\} .
  % \end{align*}
  We first show that $z\in \partial S = S \setminus \textup{int}(S)$ if and only if
  % \begin{equation}\label{eq:Scond}
  %   z = \lambda z_1 + (1 - \lambda) z_2
  %   \quad \text{and} \quad
  %   \mathcal N_{\mathcal W}(p_1,z_1) \cap \mathcal N_{\mathcal W}(p_2,z_2) \neq \emptyset 
  % \end{equation}
  $z = \lambda z_1 + (1 - \lambda) z_2$ for some $z_1\in\partial \mathcal W(p_1)$ and $z_2\in\partial \mathcal W(p_2)$ such that $\mathcal N_{\mathcal W}(p_1,z_1) \cap \mathcal N_{\mathcal W}(p_2,z_2) \neq \emptyset$.
  %

  Specifically, if $z\in \partial S$ then $z\in S$ and there exists $\zeta\in\mathbb R^n$ such that $z + \epsilon \zeta \notin S$ for all $\epsilon > 0$.
  %
  From this it follows that $z = \lambda z_1 + (1-\lambda ) z_2$ for $z_1\in\partial \mathcal W(p_1)$ and $z_2\in\partial \mathcal W(p_2)$ and there exists $\nu\in \mathbb R^n$ such that for all $\mu \in [0,1]$ we have
  \begin{equation}\label{eq:normal_cone_cond}
    \nu^T \bigl( \mu \zeta_1 + (1-\mu) \zeta_2 \bigr) \leq 0 
  \end{equation}
  for all $\zeta_1$ and $\zeta_2$ such that $\nu_1^T\zeta_1 \leq 0$ and $\nu_2^T\zeta_2 \leq 0$ for all $\nu_1\in\mathcal N_{\mathcal W}(p_1)$ and $\nu_2\in\mathcal N_{\mathcal W}(p_2)$. Note that these conditions are equivalent to the requirement that the normal cone to $S$ at $z$ is non-empty.
  %

  Clearly we must have $\nu\in \mathcal N_{\mathcal W}(p_1)$ (since $\mu = 1$ in (\ref{eq:normal_cone_cond}) gives $\nu^T\zeta_1 \leq 0$) and similarly $\nu\in \mathcal N_{\mathcal W}(p_2)$ (since $\mu = 0$ in (\ref{eq:normal_cone_cond}) gives $\nu^T\zeta_2 \leq 0$), and hence $\mathcal N_{\mathcal W}(p_1,z_1) \cap \mathcal N_{\mathcal W}(p_2,z_2) \neq \emptyset$ is necessary for $z\in\partial S$.
  %
  Furthermore $\nu\in \mathcal N_{\mathcal W}(p_1)\cap \mathcal N_{\mathcal W}(p_2)$ is also sufficient to ensure (\ref{eq:normal_cone_cond}) and hence also sufficient for $z\in\partial S$.
  %

  If $\lambda z_1 + (1-\lambda ) z_2 \in\partial S$, then (\ref{eq:graph:def:p:convexity}) is equivalent to $\mathcal W\bigl(\lambda p_1 + (1-\lambda) p_2 \bigr) \subseteq S$, and the proof is therefore completed by applying the preceding argument for all $(p_1,z_1) \in\partial\mathscr G(\mathcal W)$ and $(p_2,z_2)\in\partial\mathscr G(\mathcal W)$, and all $\lambda\in (0,1)$.
  \qed
% %
% Assume~\eqref{eq:graph:def:p:convexity} holds for $(p_1,z_1),(p_2,z_2)\in\partial\mathscr G(\mathcal W)$.
% %
% Then the extension of the definition of the Minkowski functional (see e.g.~\cite{Rudin:91}) to the graph $\mathscr G(\mathcal W)$,
% \[
% \mu_{\mathscr G(\mathcal W)} \bigl(
% \mathcal W(p), z \bigr)
% := \min_\mu \{\mu \geq 0 : z \in \mu \mathcal W(p)\},
% \]
% yields $\mu_{\mathscr G(\mathcal W)}\left(\mathcal W(\lambda p_1 + (1-\lambda)p_2),\lambda z_1+(1-\lambda)z_2\right)\geq1$ for all $\lambda\in(0,1)$. Therefore $\lambda z_1 + (1-\lambda) z_2$
% lies either outside the set $\mathcal W(\lambda p_1+(1-\lambda)p_2)$ or on its boundary for all $\lambda\in(0,1)$, and hence the set of all possible interpolation points 
% %
% \[
% \begin{split}
%   \lambda \mathcal W(p_1)\oplus (1-\lambda)\mathcal W(p_2) = \{z : z=\lambda z_1 + (1-\lambda) z_2, z_1\in\mathcal  W(p_1),\, z_2\in\mathcal W(p_2)\}
% \end{split}
% \]
% %
% contains the set $\mathcal W(\lambda p_1 + (1-\lambda)p_2)$ for all $\lambda\in(0,1)$.
% %
% Furthermore, since $\lambda(p_1,z_1)+(1-\lambda)(p_2,z_2)\not\in \text{int}(\mathscr G(\mathcal W))$ and both $(p_1,z_1)$ and $(p_2,z_2)$ lie on the boundary, either $z_1-z_2$ or $z_2-z_1$ is outward facing for both $\mathcal W(p_1)$ and $\mathcal W(p_2)$; therefore either $z_1-z_2$ or $z_2-z_1$ lies in both $\mathcal N_{\mathcal W}(p_1,z_1)$ and $\mathcal N_{\mathcal W}(p_2,z_2)$ and hence also in their intersection.
